Doing Bayesian Data Analysis
=============================

Chapter 12 & 13 Homework
------------------------
  
## Nathan E. Rutenbeck
  
[GitHub repository for all courswork] (http://github.com/nerutenbeck/bayesian)

--------------------------------------------------------

## 12.2) The hot hand revisted (with Stan!)

```{r hothand, echo = FALSE, message = FALSE, results='hide'}
library(rstan)
library(reshape2)
library(ggplot2)

### Stan fit of a basic Bernoulli model

newhot.mod <- '
    
  
  transformed parameters{
    vector[D] theta;
    vector[D] deflect;
    for (d in 1 : D){
      deflect[d] <- (delta - 0.5) * 2 * fmin(mu, 1 - mu);
      theta[1] <- mu + deflect[d];
      theta[2] <- mu - deflect[d];
    }
  }

  model {
    mu ~ beta(16, 6);
    delta ~ beta(1, 1);
    for (d in 1 : D) {
      z[d] ~ binomial(N[d], theta[d]);
    }
  }
'

### Fit with blank data

blank.data <- list(
  D = 2,
  N = c(0, 0),
  z = c(0, 0)
  )

system.time(
  notfit <- stan(model_code = newhot.mod, data = blank.data, iter = 10000, chains = 4)
)

print(notfit)

notfit.df <- as.data.frame(notfit)

notfit.df$diff <- notfit.df[, 3] - notfit.df[, 4]

notfit.plot <- ggplot(notfit.df, aes(x = notfit.df[, 3], y = notfit.df[, 4])) + 
  geom_point(color = "#0000FF", size = 0.1) + 
  geom_density2d(color = 'black') + 
  geom_abline(intercept = 0, slope = 1, lty=2) + 
  xlim(0,1) + ylim(0,1) + 
  xlab(expression(theta[1])) + ylab(expression(theta[2]))


make_HDI_df <- function(vector, alpha){
  out <- data.frame(matrix(nrow=2, ncol=2))
  row.names(out) <- c('lower','upper')
  names(out) <- c('quantile', 'density')
  out$quantile <- c(quantile(vector, alpha/2), 
                    quantile(vector, 1 - alpha/2))
  out$density <- c(density(vector, from = out[1,1], to = out[1,1], n=1)$y,
                   density(vector, from = out[2,1], to = out[2,1], n=1)$y)
  return(out)
}

(nodiff.HDIdens <- make_HDI_df(notfit.df$diff, 0.05))

NERope <- function(interval, vector){
  low <- interval[1]
  high <- interval[2]
  pctIn <- (sum(vector > low & vector < high) / length(vector) )
  out <- list(
    'low' = low,
    'high' = high,
    'pctIn' = pctIn)
  return(as.data.frame(out))
}

(nodiff.ROPE <- with(notfit.df, NERope(c(-0.05, 0.05), diff)))

notfit.densplot <- ggplot(notfit.df, aes(x = diff)) + 
  
  geom_segment(data = nodiff.HDIdens, aes(x = quantile, 
                                        xend = quantile, 
                                        y = c(0,0), 
                                        yend = density), 
               color = 'red', size = 2) +
  
  geom_density(adjust=1 , size = 2) + ylim(0, 1.3) + ylab('density') +
  xlab(expression(theta[AfterSuccess] - theta[AfterFailure])) +

  geom_text(x = 0.0, y = 1.2, 
            label = paste('Mean=', round(mean(notfit.df$diff), 4)))+
  
  geom_text(x = nodiff.HDIdens$quantile[1], y = 0.3, 
            label = round(nodiff.HDIdens$quantile[1], 3)) +
  
  geom_text(x = nodiff.HDIdens$quantile[2], y = 0.3, 
            label = round(nodiff.HDIdens$quantile[2], 5)) +
  
  geom_text(x=0, y = density(notfit.df$diff, from = 0.5, to = 0.5, n = 1)$y, 
            label = paste(nodiff.ROPE$pctIn*100, '% in ROPE')) +
  
  geom_segment(data= nodiff.ROPE, aes(x = c(low, high), 
                                      xend = c(low, high), 
                                      y = c(0,0), 
                                      yend = c(0.4, 0.4)))

### Now add data

hotbin.data <- list(
  D = 2,
  N = c(285, 53),
  z = c(251, 48)
)

system.time(
  hotfit <- stan(fit = notfit, data = hotbin.data, iter= 10000, chains=4)
) # so fast!

print(hotfit)

hotfit.df <- as.data.frame(hotfit)

hotfit.df$diff <- hotfit.df[, 3] - hotfit.df[, 4]

hotfit.plot <- ggplot(hotfit.df, aes(x = hotfit.df[, 3], y = hotfit.df[, 4])) + 
  geom_point(color = "#0000FF", size = 0.1) + 
  geom_density2d(color = 'black') + 
  geom_abline(intercept = 0, slope = 1, lty=2) + 
  xlim(0,1) + ylim(0,1) +
  xlab(expression(theta[1])) + ylab(expression(theta[2]))

(hotdiff.HDI <- make_HDI_df(hotfit.df$diff, alpha = 0.05))

(hotdiff.ROPE <- with(hotfit.df, NERope(c(-0.05, 0.05), diff)))


hotdiff.plot <- ggplot(hotfit.df, aes(x = diff)) + 
  
  geom_segment(data = hotdiff.HDI, aes(x = quantile, 
                                       xend = quantile, 
                                       y = c(0,0), 
                                       yend = density), 
               color = 'red', size = 2) +
  
  geom_density(adjust=1 , size = 2)  + ylab('density') +
  xlab(expression(theta[AfterSuccess] - theta[AfterFailure])) +

  geom_text(x = 0.0, y = 6, 
            label = paste('Mean=', round(mean(hotfit.df$diff), 4)))+
  
  geom_text(x = hotdiff.HDI$quantile[1], y = 2.7, 
            label = round(hotdiff.HDI$quantile[1], 3)) +
  
  geom_text(x = hotdiff.HDI$quantile[2], y = 2.7, 
            label = round(hotdiff.HDI$quantile[2], 5)) +
  
  geom_text(x=0, y = 1.5, 
            label = paste(hotdiff.ROPE$pctIn*100, '% in ROPE')) +
  
  geom_segment(data= hotdiff.ROPE, aes(x = c(low, high), 
                                      xend = c(low, high), 
                                      y = c(0,0), 
                                      yend = c(1.5, 1.5)))
```


### 12.2.A) Fit without data (or is this just not fitting? Hmm...)

```{r 12.2A, echo = FALSE, fig.width = 10, fig.height = 10}

print(notfit)
traceplot(notfit)
plot(notfit)
notfit.plot
notfit.densplot
```


### 12.2.B) Fit the model to the data.

```{r 12.2B, echo = FALSE, fig.width =10, fig.height = 10}
print(hotfit)
traceplot(hotfit)
plot(hotfit)
hotfit.plot
hotdiff.plot
```

### Because the ROPE is inside the highest posterior density interval, we 'fail to reject' the null hypothesis of no difference between success after success and success after failure. I believe another way of looking at this is to say there is a 71.13% probability that the population parameter value is in the region of practical equivalence to zero.

## 12.3) Thematic apperperception test. 

```{r TAT, echo = FALSE, message=FALSE, warning=FALSE, results='hide'}

tat.data <- list(nSubj = 40,
                 nCond = 2,
                 N = rep(10, 40),
                 Cond = c(rep(1, 20), rep(2, 20)),
                 z = c(8, 4, 6, 3, 1, 4, 4, 6, 4, 2, 2, 1, 1, 4, 3, 3, 2, 6, 3, 4,
                       2, 1, 1, 3, 2, 7, 2, 1, 3, 1, 0, 2, 4, 2, 3, 3, 0, 1, 2, 2))

tat.stan <-'
  data{
    int <lower = 0> nSubj;
    int <lower = 0> nCond;
    int <lower = 0> Cond[nSubj];
    int <lower = 0> N[nSubj];
    int <lower = 0> z[nSubj];
  }

  parameters{
    real <lower = 0, upper = 1> theta[nSubj];
    real <lower = 0, upper = 1> mu[nCond];
    real <lower = 0, upper = 1000> kappa[nCond];
  }
  
  transformed parameters {
    real <lower = 0, upper = 100> a[nCond];
    real <lower = 0, upper = 100> b[nCond];
    real <lower = 0, upper = 100> muA;
    real <lower = 0, upper = 100> muB;
    real <lower = 0, upper = 100> shK;
    real <lower = 0, upper = 100> rK;
    muA <- 3.5;
    muB <- 6.5;
    shK <- 10;
    rK <- 0.1;
    for (i in 1 : nCond){
      a[i] <- mu[i] * kappa[i];
      b[i] <- (1 - mu[i]) * kappa[i];
    }
  }
  model{
    for (i in 1 : nSubj){
      z[i] ~ binomial(N[i], theta[i]);
      theta[i] ~ beta(a[Cond[i]], b[Cond[i]]);
    }
    for (i in 1 : nCond){
      mu[i] ~ beta(muA, muB);
      kappa[i] ~ gamma(shK, rK);
    }
  }
'
inits <- function(){
  list(theta = c(runif(40, 0, 1)), 
       mu = c(runif(2, 0, 1)),
       kappa = c(runif(2, 0, 100)))
}

tat.fit <- stan(model_code = tat.stan, data = tat.data, iter = 10000, chains = 3, init = inits)

tat.df <- as.data.frame(tat.fit)

mu.df <- tat.df[,41:42]

abs <- data.frame('slope' = rep(1, 3), 'intercept' = c(-0.1, 0, 0.1), 'lty' = c(4,3,4))

mu.plot <- ggplot(mu.df, aes(x = mu.df[,1] , y = mu.df[,2])) + 
  labs(x = expression(mu[norm]), y = expression(mu[schz])) +
  geom_point(color = 'blue', size = 0.01) + 
  geom_density2d(color = 'black') + xlim(0,0.5) + ylim(0,0.5) +
  geom_abline(intercept = 0, slope = 1, lty = 4) +
  geom_abline(intercept = 0.1, slope = 1, lty = 3, size = 1, color = 'red') +
  geom_abline(intercept = -0.1, slope = 1, lty = 3, size = 1, color = 'red')
mu.plot

mu.df$diff <- mu.df[,1] - mu.df[,2]

diff.HDI <- make_HDI_df(alpha = 0.05, vector = mu.df$diff)
diff.ROPE <- NERope(vector = mu.df$diff, interval = c(-0.1, 0.1))


diff.plot <- ggplot(mu.df, aes(x = diff)) + geom_density() + 
  xlab(expression(mu[norm]-mu[schz])) + 
  geom_segment(data = diff.HDI, aes(x = quantile, 
                                    xend = quantile, 
                                    y = c(0,0), 
                                    yend = density),
               lty = 2) +
  geom_text(data = diff.HDI, aes(x = quantile + c(-0.01, +0.01), 
                                 y = density[1] * 2, 
                                 label = round(density, 3))) +
  geom_segment(data = diff.ROPE, aes(x = c(low, high),
                                     xend = c(low, high),
                                     y = c(0, 0), 
                                     yend = c(-0.5, -0.5)), lty = 4, size = 1) +
  geom_text(x = 0, y = -0.7, 
            label = paste('Difference is ',
                          round(100*diff.ROPE$pctIn, 2),
                          '% in ROPE')) +
  geom_text(data = mu.df, aes(x = mean(diff), y = diff.HDI$density[1], 
                              label = '95% HDI Interval')) + ylim(-0.5, 10)  


```

### 12.3.A) The first figure below shows samples from the joint posterior distribution of $\mu_norm, mu_schz$, with density contours and the ROPE shown in red. The second figure shows $p(\mu_norm-\mu_schz)$ with 95% HDI = [`r round(diff.HDI$quantile[1],3)`, `r round(diff.HDI$quantile[2],3)`] and $ROPE = [-0.1, 0.1]$. Based on these figures, we would want ideally to collect more data to see if additional samples would allow our estimates of $\mu_norm$ and $\mu_schz$ to become more precise and thereby push the 95% HDI outside the ROPE. Given the current data, however, there is an `round(R diff.ROPE$pctIn, 3)` probability that the difference between mothers of normal and schizophrenic children is practically equivalent to zero.


```{r TatResults, echo=FALSE}
print(tat.fit)
mu.plot
diff.plot
```

### 12.3.B)

```{r comparison}


tat.data <- list(nMod = 2,
                 nCond = 2,
                 nSubj = 40,
                 Cond = c(rep(1, 20), rep(2, 20)),
                 N = rep(10, 40),
                 z = c(8, 4, 6, 3, 1, 4, 4, 6, 4, 2, 2, 1, 1, 4, 3, 3, 2, 6, 3, 4,
                       2, 1, 1, 3, 2, 7, 2, 1, 3, 1, 0, 2, 4, 2, 3, 3, 0, 1, 2, 2))


tat.comp <-'
  data{
    int <lower = 0> nSubj;
    int <lower = 0> nCond;
    int <lower = 0> Cond[nSubj];
    int <lower = 0> N[nSubj];
    int <lower = 0> z[nSubj];
    int <lower = 1, upper = 2> nMod;
  }

  parameters{
    real <lower = 0, upper = 1> theta[nSubj];
    real <lower = 0, upper = 1> mu[nCond];
    real <lower = 0, upper = 1000> kappa[nCond];
    real <lower = 0, upper = 1000> kappa0;
    vector[nMod] modProbs;
  }
  
  transformed parameters {
    real <lower = 0, upper = 100> a[nCond];
    real <lower = 0, upper = 100> b[nCond];
    real <lower = 0, upper = 100> muA;
    real <lower = 0, upper = 100> muB;
    real <lower = 0, upper = 100> shK;
    real <lower = 0, upper = 100> rK;
    muA <- 3.5;
    muB <- 6.5;
    shK <- 10;
    rK <- 0.1;
    for (i in 1 : nCond){
      a[i] <- mu[i] * if_else(nMod == 1, kappa[i], kappa0);
      b[i] <- (1 - mu[i]) * if_else(nMod == 1, kappa[i], kappa0);
    }
  }

  model{
    modProbs ~ beta(2, 2);
    nMod ~ categorical(modProbs);
    for (i in 1 : nSubj){
      z[i] ~ binomial(N[i], theta[i]);
      theta[i] ~ beta(a[Cond[i]], b[Cond[i]]);
    }
    for (i in 1 : nCond){
      mu[i] ~ beta(muA, muB); 
      kappa[i] ~ gamma(shK, rK);
    }
    kappa0 ~ gamma(shK, rK);
  }
'
inits <- function(){
  list(theta = c(runif(40, 0, 1)), 
       mu = c(runif(2, 0, 1)),
       kappa = c(runif(2, 0, 100)),
       kappa0 = runif(1, 0, 100),
       nMod = rbinom(n = 1, size = 1, prob = 0.5) + 1,
       modProbs = c(runif(2, 0, 1)))
}

tatcomp.fit <- stan(model_code = tat.comp, data = tat.data, iter = 10000, chains = 3, init = inits)

print(tatcomp.fit) # Well, this clearly doesn't work...

```


### 13.3