Chapter 3
========================================================

## Nathan E. Rutenbeck
### September 10, 2013
[GitHub repository for all courswork] (http://github.com/nerutenbeck/bayesian)

--------------------------------------------------------

```{r Libraries,message=FALSE}
require(ggplot2)
require(reshape2)
```


### 3.1) Modify the coin flipping program such that the coin has $p(H)=0.8$. Change height of reference line.

#### Here is the one trial:

```{r 3.1,cache=TRUE,message=FALSE}
flipper<-function(N,p){ # make a function for whatever probability we like
  df<-data.frame(n=c(1:N), # vector from 1:N
                 trial=sample(x=c(0,1), prob=c(1-p,p), size=N,replace=T)) # Define Boolean sample
  df$r<-cumsum(df$trial)
  df$runprop<-df$r/df$n
  plot<-ggplot(df,aes(x=n,y=runprop))+geom_point()+
    xlim(c(1,N))+ylim(c(0,1))+xlab('Flip number')+ylab('Proportion of Heads')+geom_abline(intercept=p,slope=0,lty=3)
  return(plot)
}
flipper(N=500,p=0.8)
```

#### And here is multiple trials just for fun...

```{r 3.1.2,cache=TRUE}
flip.dist<-function(n.sims,N,p){
  tmp<-data.frame(flip=c(1:N))
  for(i in 1:n.sims){
    tmp[,i+1]<-(cumsum(sample(x=c(0,1),prob=c(1-p,p),size=N,replace=T))/c(1:N))
  }
  df<-melt(tmp,id='flip')
  plot<-ggplot(df,aes(x=flip,y=value))+geom_point(cex=0.8)+
    geom_abline(intercept=p,slope=0,color='yellow')+xlim(c(0,N))+ylim(c(0,1))+
    xlab('Proportion of Heads')+ylab('Flip number')
  return(plot)
}
flip.dist(n.sims=1000,N=500,p=0.8) 
```

### 3.4) Use a normal curve to describe beliefs.

#### 3.4.A) First show the normal curve with the indefinite integral approximated numerically

```{r 3.4.A.1,fig.width=7,message=FALSE}
mu=0 # Define mean
sigma=0.2 # Define standard deviation
xlow=mu-(3*sigma) # Define min for display
xhigh=mu+(3*sigma) # Define max for display
dx=0.02 # Set an arbitrary but low dx
df<-data.frame(x=seq(from=xlow,to=xhigh,by=dx)) # I like to work with dataframes when possible. Define x values
df$y=(1/(sigma*sqrt(2*pi)))*exp(-.5 * ((df$x-mu)/sigma)^2) # Define y values
area=sum(dx*df$y) # Approximate the indefinite integral

plot<-ggplot(df,aes(x=x,y=y))+geom_line()+ggtitle('Normal PDF: mu=0, sigma=0.2') # Build the plot
plot+geom_segment(data=df,aes(x=x,y=0,xend=x,yend=y))+geom_text(y=1.95,x=-0.33,label=paste('Area approximates to',signif(area,3))) # Display the normal curve plus indefinite integral
```

#### Next show the normal curve with the definite integral from $x-\sigma$ to $x+\sigma$ approximated numerically

```{r3.4.A.2, fig.width=7}
xlow2=mu-sigma 
xhigh2=mu+sigma
df2<-data.frame(x=seq(from=xlow2,to=xhigh2,by=dx)) # I like to work with dataframes when possible. Define x values
df2$y=(1/(sigma*sqrt(2*pi)))*exp(-.5 * ((df2$x-mu)/sigma)^2) # Define y values
area2=sum(dx*df2$y)-dx # Approximate the definite integral between mu-sigma and mu+sigma

plot2<-plot+geom_segment(data=df2,aes(x=x,y=0,xend=x,yend=y)) # Build the plot
plot2+geom_text(y=1.95,x=-0.35,label=paste('Area approximates to',signif(area2,3))) # Display the normal curve plus definite integral
```

### 3.4.B) I guess I should just display what this curve looks like and the area below it? The curve below shows the area within the approximately 68% HDI for this belief about the heights of women.

```{r 3.4.B, fig.width=7}
mu=162 # Define mean
sigma=162-147 # Define standard deviation
xlow=mu-(3*sigma) # Define min for display
xhigh=mu+(3*sigma) # Define max for display
dx=0.01 # Set an arbitrary but low dx
df<-data.frame(x=seq(from=xlow,to=xhigh,by=dx))
df$y=(1/(sigma*sqrt(2*pi)))*exp(-.5 * ((df$x-mu)/sigma)^2) # Define y values
area=sum(dx*df$y) # Approximate the indefinite integral

plot<-ggplot(df,aes(x=x,y=y))+geom_line()+ggtitle(paste('Normal PDF: mu=',mu, ', sigma=',sigma)) # Build the plot
plot+geom_segment(data=df[df$x>(mu-sigma)&df$x<(mu+sigma),],aes(x=x,y=0,xend=x,yend=y))+
  geom_text(y=0.025,x=135,label='Area approximates to 0.68') # Display the normal curve plus indefinite integral
```