Doing Bayesian Data Analysis
Chapter 6 Homework
========================================================

## Nathan E. Rutenbeck
### October 1, 2013
[GitHub repository for all courswork] (http://github.com/nerutenbeck/bayesian)

--------------------------------------------------------

```{r load functions}
rm(list=ls())
source('G:/documents/coursework/bayesian/functions/BernGrid.R')
```


### 6.2) Practice specifying a non-beta prior. Specify prior, display posterior beliefs
 
```{r 6.2}
pTheta<-c(50:1,rep(1,50),1:50,50:1,rep(1,50),1:50)
pTheta<-pTheta/sum(pTheta)
plot(pTheta,type='l')
width<-1/length(pTheta)
Theta<-seq(from=width/2,to=1-width/2,by=width)
data<-c(rep(1,15),rep(0,5))
BernGrid(Theta=Theta,pTheta=pTheta,Data=data)
```

### 6.3) Use BernGrid for sequential update.

```{r 6.3}
data<-c(0,1,1,1) # Four trials, 3 heads
(t1<-BernGrid(Theta=Theta,pTheta=pTheta,Data=data))
data2<-c(rep(1,12),rep(0,4))
t2<-BernGrid(Theta=Theta,pTheta=t1,Data=data2)
```

### 6.4) Connect HDIs to the real world.

#### 6.4.A) Prior belief is neutral uniform distribution (0.5). Let $p(1)=p(A), p(0)=p(B)$.

```{r 6.4}
Theta<-seq(from=0,to=1,by=0.01)
pTheta<-c(rep(1/101,length(Theta)))
data<-c(rep(1,58),rep(0,42))
(t1<-BernGrid(Theta=Theta,pTheta=pTheta,Data=data))
```


#### 6.4.B) It is still credible to believe (based on 95% HDI) that the population is equally divided.

#### 6.4.C) Update beliefs with new data.

```{r 6.4.C}
data2<-c(rep(1,57),rep(0,43))
(t2<-BernGrid(Theta=Theta,pTheta=t1,Data=data2))
```

#### 6.4.D) It is no longer credible to believe that the population is equally divided.

### 6.7) Use a grid approximation to compare models
```{r 6.7}
theta<-seq(from=0,to=1,by=0.01)
pTheta1<-c(theta^2/sum(theta^2))
pTheta2<-c((1-theta)^2/sum((1-theta)^2))
data<-c(rep(1,6),rep(0,2))
mod1<-BernGrid(Theta=theta,pTheta=pTheta1,Data=data)
mod2<-BernGrid(Theta=theta,pTheta=pTheta2,Data=data)
(BF<-0.00597/0.00128)
```
#### Bayes factor of Model1:Model2 = 4.664. Model 1 appears better,